---
title: "Cross-Validation"
author: "Yiming Chen"
date: "2025-11-14"
output: html_document
---
```{r}
# load packages
library(GDINA)
library(mirt)
library(mvtnorm)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
```

# Split simualted item response data into k-folds for cross-validation
Each entry is designed to have missing entries reserved for test set.

## A function to randomly assign fold numbers for each row (individual)
```{r}
gen_data_for_row <- function(J,nfold){

  # Number of complete repeats per fold
  base_repeats <- floor(J / nfold)
  remaining <- J %% nfold

  # Create deterministic part
  folds <- rep(1:nfold, times = base_repeats)

  # Add randomly selected leftover folds
  if (remaining > 0) {
    folds <- c(folds, sample(1:nfold, size = remaining))
  }

  # Shuffle the fold assignments for this row
  sample(folds, size = J)
}
```

```{r}
# example use case
# each run time produces different results -> should set.seed()
gen_data_for_row(30, 5)
gen_data_for_row(30, 10)
```

## A function to split item response data into k-folds 
```{r}
split_data <- function(data,kfold=5,prop=NULL){
  if(!is.matrix(data)){
    data <- as.matrix(data)
  }

  N <- nrow(data)
  J <- ncol(data)
  ind <- matrix(NA,nrow = N,ncol = J)
  for(i in 1:N){
    ind[i,] <- gen_data_for_row(J,kfold)
  }
  newdata <- list()
  for(n in 1:kfold){
    kfold.loc <- which(ind==n)
    temp <- data
    temp[kfold.loc] <- NA
    newdata[[n]] <- as.data.frame(temp)
  }

  ret <- list(cvdata=newdata,cvfold=ind,kfold=kfold,rawdata=data)
}
```

# Cross Validated Congitive Diagnostic Model Fitting
```{r}
cv_fit_cdm_mp <- function(cvdata, Q, estimator="mp", ...){
  N <- nrow(cvdata$rawdata)
  J <- ncol(cvdata$rawdata)
  K <- ncol(Q)
  ind <- cvdata$cvfold

  metrics <- fit <- obs_pred_p <- train_metrics <- train_obs_pred_p <- list()
  for(n in 1:cvdata$kfold){
    nfold.loc <- which(ind==n)
    train.loc <- which(ind!=n)
    nfold.length <- length(nfold.loc)
    train.length <- length(train.loc)
    
    # Fit model
    fit[[n]] <- GDINA(cvdata$cvdata[[n]], Q, ...)

    # Validation sample metrics
    person_parm <- personparm(fit[[n]], what = estimator)
    item_parm <- coef(fit[[n]], what = "delta")
    pred_p <- NULL
    
    for (j in 1:J) {
      qj <- Q[j, ]
      pred_p <- cbind(pred_p, plogis(cbind(1, person_parm[, which(qj == 1), drop = FALSE]) %*% item_parm[[j]]))
    }
    
    pred_Y <- 1 * (pred_p > matrix(runif(N * J), nrow = N, ncol = J))
    
    p <- pred_p[nfold.loc]
    pred.data <- pred_Y[nfold.loc]
    
    obs.data <- cvdata$rawdata[nfold.loc]
    opp <- data.frame(obs.data, pred.data, p)
    obs_pred_p[[n]] <- opp[complete.cases(opp), ]
    metrics[[n]] <- classification_metrics(obs_pred_p[[n]]$obs.data, 
                                           obs_pred_p[[n]]$pred.data, 
                                           obs_pred_p[[n]]$p)
    
    # Training sample metrics
    train_p <- pred_p[train.loc]
    train_pred.data <- pred_Y[train.loc]
    
    train_obs.data <- cvdata$rawdata[train.loc]
    train_opp <- data.frame(train_obs.data, train_pred.data, train_p)
    train_obs_pred_p[[n]] <- train_opp[complete.cases(train_opp), ]
    train_metrics[[n]] <- classification_metrics(train_obs_pred_p[[n]]$train_obs.data, 
                                                 train_obs_pred_p[[n]]$train_pred.data, 
                                                 train_obs_pred_p[[n]]$train_p)
  }

  output <- combine_metrics_table(metrics)
  train_output <- combine_metrics_table(train_metrics)
  
  list(
    metrics = output,
    obs_pred_p = obs_pred_p,
    metrics_kfold = metrics,
    train_metrics = train_output,
    train_obs_pred_p = train_obs_pred_p,
    train_metrics_kfold = train_metrics
  )
}
```

```{r}
cv_fit_cdm_attr <- function(cvdata,Q,estimator="MAP",...){
  N <- nrow(cvdata$rawdata)
  J <- ncol(cvdata$rawdata)
  K <- ncol(Q)
  ind <- cvdata$cvfold

  metrics <- fit <- obs_pred_p <- train_metrics <- train_obs_pred_p <- list()
  for(n in 1:cvdata$kfold){
    nfold.loc <- which(ind==n)
    train.loc <- which(ind!=n)
    nfold.length <- length(nfold.loc)
    train.length <- length(train.loc)

    # Fit model
    fit[[n]] <- GDINA(cvdata$cvdata[[n]], Q, ...)

    # Validation sample metrics
    profiles <- as.matrix(personparm(fit[[n]],what=estimator)[,1:K])
    gr <- GDINA:::matchMatrix(as.matrix(attributepattern(K)),profiles)
    cdmp <- t(fit[[n]]$LC.prob[,gr])
    p <- cdmp[nfold.loc]
    pred.data <- rbinom(nfold.length,1,p)
    obs.data <- cvdata$rawdata[nfold.loc]
    opp <- data.frame(obs.data,pred.data,p)
    obs_pred_p[[n]] <- opp[complete.cases(opp),]
    metrics[[n]] <- classification_metrics(obs_pred_p[[n]]$obs.data,
                                           obs_pred_p[[n]]$pred.data,
                                           obs_pred_p[[n]]$p)

    # Training sample metrics
    train_p <- cdmp[train.loc]
    train_pred.data <- rbinom(train.length,1,train_p)
    train_obs.data <- cvdata$rawdata[train.loc]
    train_opp <- data.frame(train_obs.data,train_pred.data,train_p)
    train_obs_pred_p[[n]] <- train_opp[complete.cases(train_opp),]
    train_metrics[[n]] <- classification_metrics(train_obs_pred_p[[n]]$train_obs.data,
                                                 train_obs_pred_p[[n]]$train_pred.data,
                                                 train_obs_pred_p[[n]]$train_p)
  }

  output <- combine_metrics_table(metrics)
  train_output <- combine_metrics_table(train_metrics)

  list(
    metrics = output,
    obs_pred_p = obs_pred_p,
    metrics_kfold = metrics,
    train_metrics = train_output,
    train_obs_pred_p = train_obs_pred_p,
    train_metrics_kfold = train_metrics
  )
}
```

```{r}
cv_fit_irt <- function(cvdata, model_specification, estimator="MAP",...){
  N <- nrow(cvdata$rawdata)
  J <- ncol(cvdata$rawdata)
  ind <- cvdata$cvfold

  metrics <- fit <- obs_pred_p <- train_metrics <- train_obs_pred_p <- list()
  for(n in 1:cvdata$kfold){
    p <- matrix(NA,N,J)
    kfold.loc <- which(ind==n)
    train.loc <- which(ind!=n)
    kfold.length <- length(kfold.loc)
    train.length <- length(train.loc)
    
    # Fit model
    fit[[n]] <- mirt::mirt(cvdata$cvdata[[n]], ...)
    
    # Validation sample metrics
    theta <- mirt::fscores(fit[[n]],method=estimator, QMC = TRUE)
    for(j in 1:J){
      p[,j] <- mirt::expected.item(mirt::extract.item(fit[[n]],item = j),theta) 
    }
    vp <- p[kfold.loc]
    pred.data <- rbinom(kfold.length,1,vp)
    obs.data <- cvdata$rawdata[kfold.loc]
    opp <- data.frame(obs.data,pred.data,vp)
    obs_pred_p[[n]] <- opp[complete.cases(opp),]
    metrics[[n]] <- classification_metrics(obs_pred_p[[n]]$obs.data,
                                           obs_pred_p[[n]]$pred.data,
                                           obs_pred_p[[n]]$vp)
    
    # Training sample metrics
    train_p <- p[train.loc]
    train_pred.data <- rbinom(train.length,1,train_p)
    train_obs.data <- cvdata$rawdata[train.loc]
    train_opp <- data.frame(train_obs.data,train_pred.data,train_p)
    train_obs_pred_p[[n]] <- train_opp[complete.cases(train_opp),]
    train_metrics[[n]] <- classification_metrics(train_obs_pred_p[[n]]$train_obs.data,
                                                 train_obs_pred_p[[n]]$train_pred.data,
                                                 train_obs_pred_p[[n]]$train_p)
  }

  output <- combine_metrics_table(metrics)
  train_output <- combine_metrics_table(train_metrics)
  
  list(
    metrics = output,
    obs_pred_p = obs_pred_p,
    metrics_kfold = metrics,
    train_metrics = train_output,
    train_obs_pred_p = train_obs_pred_p,
    train_metrics_kfold = train_metrics
  )
}
```

```{r}
classification_metrics <- function(observed, predicted, p) {
  # Ensure input vectors are numeric
  observed <- as.numeric(as.character(observed))
  predicted <- as.numeric(as.character(predicted))
  p <- as.numeric(p)

  # Confusion matrix components
  tp <- sum(predicted == 1 & observed == 1)
  tn <- sum(predicted == 0 & observed == 0)
  fp <- sum(predicted == 1 & observed == 0)
  fn <- sum(predicted == 0 & observed == 1)
  cat("\nTP=", tp, " TN=", tn, " FP=", fp, "FN=", fn)
  
  # Basic metrics
  accuracy <- mean(predicted == observed)
  precision <- if ((tp + fp) == 0) NA else tp / (tp + fp) # PPV 
  recall <- if ((tp + fn) == 0) NA else tp / (tp + fn) # sensitivity
  specificity <- if ((tn + fp) == 0) NA else tn / (tn + fp)
  f1_score <- if (is.na(precision) || is.na(recall) || (precision + recall == 0)) NA else 2 * precision * recall / (precision + recall)

  # MCC
  mcc_denom <- sqrt(
    as.numeric((tp + fp)) *
    as.numeric((tp + fn)) *
    as.numeric((tn + fp)) *
    as.numeric((tn + fn))
  )
  mcc <- if (mcc_denom == 0) NA else ((tp * tn) - (fp * fn)) / mcc_denom

  # AUC - TPR (recall, Y-axis) vs. FPR (X-axis) Curve
  auc_val <- tryCatch({
    pROC::auc(pROC::roc(observed, p))
  }, error = function(e) NA)

  # PR AUC - Precision (Y-axis) - Recall (X-axis) Curve
  pr_auc_val <- tryCatch({
    fg <- p[observed == 1] # foreground
    bg <- p[observed == 0] # background
    # class0 belongs to the positive class -> observed == 1 -> fg
    # class1 belongs to the negative class -> observed == 0 -> bg
    pr <- PRROC::pr.curve(scores.class0 = fg, scores.class1 = bg)
    pr$auc.integral
  }, error = function(e) NA)

  # Log Loss
  log_loss <- {
    eps <- 1e-15
    p <- pmin(pmax(p, eps), 1 - eps)
    -mean(observed * log(p) + (1 - observed) * log(1 - p))
  }

  # Confusion matrix
  confusion_matrix <- matrix(c(tn, fp, fn, tp), nrow = 2, byrow = TRUE,
                             dimnames = list("Predicted" = c("0", "1"),
                                             "Actual" = c("0", "1")))

  # Return all
  return(list(
    accuracy = accuracy,
    precision = precision,
    recall = recall,
    specificity = specificity,
    f1_score = f1_score,
    mcc = mcc,
    auc = auc_val,
    pr_auc = pr_auc_val,
    log_loss = log_loss,
    confusion_matrix = confusion_matrix
  ))
}
```

```{r}
combine_metrics_table <- function(metrics_list,digits=3) {
  # Extract all scalar metric names (exclude confusion_matrix)
  metric_names <- setdiff(names(metrics_list[[1]]), "confusion_matrix")

  # Convert each list element to a one-row data frame
  metrics_df <- do.call(rbind, lapply(metrics_list, function(m) {
    as.data.frame(as.list(m[metric_names]))
  }))

  # Compute mean for each metric
  average_row <- colMeans(metrics_df, na.rm = TRUE)

  # Add the average as the final row
  metrics_df <- rbind(metrics_df, Average = average_row)

  metrics_df <- round(metrics_df, digits)
  # Set row names: Fold_1, Fold_2, ..., Average
  rownames(metrics_df) <- c(paste0("Fold_", seq_len(nrow(metrics_df) - 1)), "Average")

  return(metrics_df)
}
```

## Define a function to run cross validation
```{r}
Q_complex_3 <- matrix(c(
            1,0,0, 0,1,0, 0,0,1,
            1,1,0, 1,0,1, 0,1,1,
            1,0,0, 0,1,0, 0,0,1,
            1,1,0, 1,0,1, 0,1,1,
            1,0,0, 0,1,0, 0,0,1,
            1,1,0, 1,0,1, 0,1,1,
            1,0,0, 0,1,0, 0,0,1,
            1,1,0, 1,0,1, 0,1,1,
            1,1,1, 1,1,1, 1,1,1,
            1,1,1, 1,1,1, 1,1,1
        ), ncol = 3, byrow = TRUE)

# for complex structure: function to align model specification to Q-matrix
to_model_string <- function(Q){
  factors <- lapply(seq_len(ncol(Q)), function(f){
    items <- which(Q[, f] == 1)
    paste0("F", f, " = ", paste(items, collapse = ","))
  })
  cov <- paste0("COV = ", paste0("F", 1:ncol(Q), collapse = "*"))
  paste(c(factors, cov), collapse = "\n")
}

# define Q-matrix in CDM / model_specification in IRT for different conditions 
run_cv_fit <- function(cvobj, name) {
  str <- if (grepl("Simple", name)) "Simple" else "Complex"
  nd <- as.numeric(sub(".*dim([0-9]+)_.*", "\\1", name))
  
  if (str == "Simple") {
    block_size <- 30 / nd
    Q <- matrix(0, nrow = 30, ncol = nd)
    for (i in 1:nd) {
      start_index <- (i - 1) * block_size + 1
      end_index <- i * block_size
      Q[start_index:end_index, i] <- 1
    }
    factor_lines <- vapply(1:nd, function(i){
      start_index <- (i - 1) * block_size + 1
      end_index <- i * block_size
      paste0("F", i, " = ", paste(start_index:end_index, collapse = ","))
    }, character(1))
    cov_line <- paste0("COV = ", paste0("F", 1:nd, collapse = "*"))
    model_specification <- paste(c(factor_lines, cov_line), collapse = "\n")
    
  } else if (str == "Complex" & nd == 3) {
    Q <- Q_complex_3
    model_specification <- to_model_string(Q)
    
  } else if (str == "Complex" & nd == 5) {
    Q <- sim30GDINA$simQ
    model_specification <- to_model_string(Q)
  }
  
  res_mp <- cv_fit_cdm_mp(cvobj, Q, 
                          model = "LLM", verbose = 0
                          #, mono.constraint = TRUE
                          )
  res_attribute <- cv_fit_cdm_attr(cvobj, Q, 
                                   model = "LLM", verbose = 0
                                   #, mono.constraint = TRUE
                                   )
  
  # Warning: MHRM terminated after 2000 iterations. --> increase iterations to 5000
  res_irt <- cv_fit_irt(cvobj, model_specification, 
                        method = "MHRM", 
                        technical = list(NCYCLES = 5000), 
                        verbose = FALSE)
  
  return(list(
    mp = res_mp, 
    attribute = res_attribute, 
    irt = res_irt)) # theta
}
```

# Apply split_data function to get CDM cvdata
```{r}
# set seed to ensure reproducibility in splitting data 
set.seed(101)

# read CDM simulated data and save them into a list called cvdata_cdm
files <- list.files("../CDM_sim/data", pattern = "^data_.*\\.csv$", full.names = TRUE)
names(files) <- basename(files)    

# cdm_data is a list of 120 data frames
cdm_data <- lapply(files, read.csv)
length(cdm_data)

# generate cvdata using split_data function
cvdata_cdm <- lapply(cdm_data, split_data, kfold = 10)

# access data name for each data
names(cvdata_cdm)[[1]]

# access data name for multiple data
names(cvdata_cdm)[1:10]
```

```{r}
dir.create("../CDM_CV_results", showWarnings = FALSE)
dir.create("../IRT_CV_results", showWarnings = FALSE)
```

# Apply run_cv_fit function to CDM data
```{r, message=FALSE, eval=FALSE}
set.seed(101)

for (i in names(cvdata_cdm)) {
  
  cvobj <- cvdata_cdm[[i]]

  res <- run_cv_fit(cvobj = cvobj, name = i)
  
  clean_i <- sub("^data_", "res_", i)
  clean_i <- sub("\\.csv$", ".rds", clean_i)
  
  # save each condition separately
  saveRDS(res, file = file.path("../CDM_CV_results", clean_i))
  
  cat("Saved:", clean_i, "\n")
}
```

```{r}
test <- readRDS("../CV_results_separate/res_Complex_cor0_dim3_rep1.rds")
test$mp$metrics
```

```{r}
pattern <- "res_Complex_cor0_dim3_rep"

files <- list.files("../CDM_CV_results", pattern = pattern, full.names = TRUE)

# Load all 10 replications
res_list <- lapply(files, readRDS)
```

```{r}
res_list[[1]]$mp$metrics$accuracy[11]
```

```{r}
extract_metrics_condition <- function(structure, cor, dim,
                                      folder = "../xxx") {
  
  # determine data source (CDM or IRT)
  source_name <- sub("_CV_results$", "", basename(folder))
  
  # build pattern for the condition
  pattern <- sprintf("^res_%s_%s_%s", structure, cor, dim)
  
  # get data for all 10 replications 
  files <- list.files(folder, pattern = pattern, full.names = TRUE)
  
  if (length(files) == 0) {
    stop("No files found for pattern: ", pattern)
  }
  
  # read all replications
  res_list <- lapply(files, readRDS)
  
  # extract replicate numbers from file names
  reps <- as.numeric(str_extract(basename(files), "(?<=rep)\\d+"))
  
  methods <- c("mp", "attribute", "irt")
  
  # extract metrics for all reps × methods
  metrics_df <- bind_rows(lapply(seq_along(res_list), function(i) {
    res <- res_list[[i]]
    replicate <- reps[i]
    
    bind_rows(lapply(methods, function(m) {
      tibble(
        data_source = source_name,
        method = m,
        structure = structure,
        correlation = as.numeric(str_extract(cor, "\\d+(\\.\\d+)?")),
        dim = as.numeric(str_extract(dim, "\\d+")),
        replicate = replicate,
        accuracy = res[[m]]$metrics$accuracy[11],
        precision = res[[m]]$metrics$precision[11],
        recall  = res[[m]]$metrics$recall[11],
        specificity  = res[[m]]$metrics$specificity[11],
        f1_score  = res[[m]]$metrics$f1_score[11],
        mcc = res[[m]]$metrics$mcc[11],
        auc  = res[[m]]$metrics$auc[11],
        pr_auc = res[[m]]$metrics$pr_auc[11],
        log_loss  = res[[m]]$metrics$log_loss[11]
      )
    }))
  }))
  
  return(metrics_df)
}
```

```{r}
extract_metrics_condition(
  structure = "Complex", 
  cor = "cor0",
  dim = "dim3",
  folder = "../CDM_CV_results"
)
```

# Apply split_data function to get IRT cvdata
```{r}
# set seed to ensure reproducibility in splitting data 
set.seed(101)

# read IRT simulated data and save them into a list called cvdata_irt
files <- list.files("../IRT_sim/data", pattern = "^data_.*\\.csv$", full.names = TRUE)
names(files) <- basename(files)    

# irt_data is a list of 120 data frames
irt_data <- lapply(files, read.csv)
length(irt_data)

# generate cvdata using split_data function
cvdata_irt <- lapply(irt_data, split_data, kfold = 10)

# access data name for each data
names(cvdata_irt)[[1]]
names(cvdata_irt)[1:10]
```

```{r, message=FALSE, eval=FALSE}
set.seed(101)

for (i in names(cvdata_irt)) {
  
  cvobj <- cvdata_irt[[i]]

  res <- run_cv_fit(cvobj = cvobj, name = i)
  
  clean_i <- sub("^data_", "res_", i)
  clean_i <- sub("\\.csv$", ".rds", clean_i)
  
  # save each condition separately
  saveRDS(res, file = file.path("../IRT_CV_results", clean_i))
  
  cat("Saved:", clean_i, "\n")
}
```

## Read in results
```{r}
structures <- c("Simple", "Complex")
cors <- c("cor0", "cor0.3", "cor0.6")  
dims <- c("dim3", "dim5")

irt_result <- bind_rows(lapply(structures, function(S) {
  bind_rows(lapply(cors, function(C) {
    bind_rows(lapply(dims, function(D) {
      extract_metrics_condition(S, C, D,
                                "../IRT_CV_results")
    }))
  }))
}))
```

```{r}
# readRDS("../CV_10fold/irt_data_res_simple_dim5.rds")$data_Simple_cor0_dim5_rep3.csv$mp$metrics
# readRDS("../CV_10fold/irt_data_res_simple_dim5.rds")$data_Simple_cor0_dim5_rep3.csv$attr$metrics
# readRDS("../CV_10fold/irt_data_res_simple_dim5.rds")$data_Simple_cor0_dim5_rep3.csv$irt$metrics
```

```{r}
# readRDS("../CV_10fold/irt_data_res_complex_dim3.rds")$data_Complex_cor0.3_dim3_rep5.csv$mp$metrics
# readRDS("../CV_10fold/irt_data_res_complex_dim3.rds")$data_Complex_cor0.3_dim3_rep5.csv$attr$metrics
# readRDS("../CV_10fold/irt_data_res_complex_dim3.rds")$data_Complex_cor0.3_dim3_rep5.csv$irt$metrics
```

```{r}
cdm_result <- bind_rows(lapply(structures, function(S) {
  bind_rows(lapply(cors, function(C) {
    bind_rows(lapply(dims, function(D) {
      extract_metrics_condition(S, C, D,
                                "../CDM_CV_results")
    }))
  }))
}))
```


```{r}
# readRDS("../CV_10fold/cdm_data_res_simple_dim5_10folds.rds")$data_Simple_cor0.3_dim5_rep10.csv$mp$metrics
# readRDS("../CV_10fold/cdm_data_res_simple_dim5_10folds.rds")$data_Simple_cor0.3_dim5_rep10.csv$attr$metrics
# readRDS("../CV_10fold/cdm_data_res_simple_dim5_10folds.rds")$data_Simple_cor0.3_dim5_rep10.csv$irt$metrics
```

```{r}
# (cdm_result <- cdm_result %>%
#   mutate(method = factor(method, levels = c("mp", "attribute", "irt"))) %>% 
#   group_by(data_source, method, structure, correlation, dim) %>%
#   summarize(
#     mean_accuracy = mean(accuracy),
#     mean_precision = mean(precision),
#     mean_recall = mean(recall),
#     mean_specificity = mean(specificity),
#     mean_f1_score = mean(f1_score),
#     mean_mcc = mean(mcc),
#     mean_auc = mean(auc),
#     mean_pr_auc = mean(pr_auc),
#     mean_log_loss = mean(log_loss),
#     .groups = "drop"
#     ) %>%
#   arrange(structure, correlation, dim, method))
```

```{r}
# (irt_result <- irt_result %>%
#   mutate(method = factor(method, levels = c("mp", "attribute", "irt"))) %>% 
#   group_by(data_source, method, structure, correlation, dim) %>%
#   summarize(
#     mean_accuracy = mean(accuracy),
#     mean_precision = mean(precision),
#     mean_recall = mean(recall),
#     mean_specificity = mean(specificity),
#     mean_f1_score = mean(f1_score),
#     mean_mcc = mean(mcc),
#     mean_auc = mean(auc),
#     mean_pr_auc = mean(pr_auc),
#     mean_log_loss = mean(log_loss),
#     .groups = "drop"
#     ) %>%
#   arrange(structure, correlation, dim, method))
```

```{r}
cv_result <- rbind(cdm_result, irt_result)
cv_result
```
```{r}
irt_result
```

## Create Difference Plots
### MP - Theta
```{r}
mp_theta <- cv_result %>%
  filter(method == "mp"| method == "irt") %>%
  select(data_source, replicate, method, structure, correlation, dim, accuracy, mcc, log_loss) %>%
  group_by(data_source, structure, correlation, dim) %>%
  pivot_wider(names_from = method, values_from = c(accuracy, mcc, log_loss)) %>%
  mutate(
  accuracy_diff = accuracy_mp - accuracy_irt,
  mcc_diff = mcc_mp - mcc_irt,
  log_loss_diff = log_loss_mp - log_loss_irt) %>%
ungroup()
```

```{r}
mp_theta %>%
  filter(data_source == "CDM" & structure == "Complex")

mp_theta %>%
  filter(data_source == "IRT" & structure == "Complex")
```

```{r}
mp_theta_acc_hlines <- expand.grid(
  structure   = c("Simple", "Complex"),
  dim         = c(3, 5),
  correlation = c(0, 0.3, 0.6)
) %>%
  mutate(
    hline = c(
      0.04, 0.03, # simple dim3 cor0, complex dim3 cor 0
      0.05, 0.06, # simple dim5 cor0, complex dim5 cor 0
      0.04, 0.02, # simple dim3 cor0.3, complex dim3 cor0.3
      0.03, 0.04, # simple dim5 cor0.3, complex dim5 cor0.3
      0.03, 0.02, # simple dim3 cor0.6, complex dim3 cor0.6
      0.01, 0.01 # simple dim5 cor0.6, complex dim5 cor0.6
    )
  )

mp_theta_logloss_hlines <- expand.grid(
  structure   = c("Simple", "Complex"),
  dim         = c(3, 5),
  correlation = c(0, 0.3, 0.6)
) %>%
  mutate(
    hline = c(
      -0.05, -0.04,
      -0.05, -0.06,
      -0.04, -0.03,
      -0.03, -0.03,
      -0.01, 0.00,
      -0.01, -0.02
    )
  )
```

```{r, fig.width=10, fig.height=6, dpi=300}
ggplot(mp_theta, aes(x = replicate, y = accuracy_diff)) +
  geom_point(shape = 21, size = 2, aes(fill = data_source)) +
  geom_hline(data = mp_theta_acc_hlines, aes(yintercept = hline),
             color = "black", linetype = "dashed", linewidth = 0.6) +
  scale_fill_manual(
    name = "Accuracy Difference",
    values = c("CDM" = "darkolivegreen1", "IRT" = "deeppink2")
  ) +
  scale_x_continuous(name = "Replication", breaks = 1:10) +
  ylab("Accuracy Difference") +
  ggtitle("Accuracy Difference by MP - Theta") +
  theme_bw() +
  facet_grid(structure + dim ~ correlation) 

ggplot(mp_theta, aes(x = replicate, y = log_loss_diff)) +
  geom_point(shape = 21, size = 2, aes(fill = data_source)) +
  geom_hline(data = mp_theta_logloss_hlines, aes(yintercept = hline),
             color = "black", linetype = "dashed", linewidth = 0.6) +
  scale_fill_manual(
    name = "Log Loss Difference",
    values = c("CDM" = "darkolivegreen1", "IRT" = "deeppink2")
  ) +
  scale_x_continuous(name = "Replication", breaks = 1:10) +
  ylab("Log Loss Difference") +
  ggtitle("Log Loss Difference by MP - Theta") +
  theme_bw() +
  facet_grid(structure + dim ~ correlation) 
```

### Attribute - Theta
```{r}
attr_theta <- cv_result %>%
  filter(method == "attribute"| method == "irt") %>%
  select(data_source, replicate, method, structure, correlation, dim, accuracy, mcc, log_loss) %>%
  group_by(data_source, structure, correlation, dim) %>%
  pivot_wider(names_from = method, values_from = c(accuracy, mcc, log_loss)) %>%
  mutate(
  accuracy_diff = accuracy_attribute - accuracy_irt,
  mcc_diff = mcc_attribute - mcc_irt,
  log_loss_diff = log_loss_attribute - log_loss_irt) %>%
ungroup()
```

```{r, fig.width=10, fig.height=6, dpi=300}
ggplot(attr_theta, aes(x = replicate, y = accuracy_diff)) +
  geom_point(shape = 21, size = 2, aes(fill = data_source)) +
  geom_hline(yintercept = 0,
             color = "black", linetype = "dashed", linewidth = 0.8) +
  scale_fill_manual(
    name = "Accuracy Difference",
    values = c("CDM" = "darkolivegreen1", "IRT" = "deeppink2")
  ) +
  scale_x_continuous(name = "Replication", breaks = 1:10) +
  ylab("Accuracy Difference") +
  ggtitle("Accuracy Difference by Attribute - Theta") +
  theme_bw() +
  facet_grid(structure + dim ~ correlation) 

ggplot(attr_theta, aes(x = replicate, y = log_loss_diff)) +
  geom_point(shape = 21, size = 2, aes(fill = data_source)) +
  geom_hline(yintercept = 0,
             color = "black", linetype = "dashed", linewidth = 0.8) +
  scale_fill_manual(
    name = "Log Loss Difference",
    values = c("CDM" = "darkolivegreen1", "IRT" = "deeppink2")
  ) +
  scale_x_continuous(name = "Replication", breaks = 1:10) +
  ylab("Log Loss Difference") +
  ggtitle("Log Loss Difference by Attribute - Theta") +
  theme_bw() +
  facet_grid(structure + dim ~ correlation) 
```

### Attribute - MP
```{r}
attr_mp <- cv_result %>%
  filter(method == "mp"| method == "attribute") %>%
  select(data_source, replicate, method, structure, correlation, dim, accuracy, mcc, log_loss) %>%
  group_by(data_source, structure, correlation, dim) %>%
  pivot_wider(names_from = method, values_from = c(accuracy, mcc, log_loss)) %>%
  mutate(
    accuracy_diff = accuracy_attribute - accuracy_mp,
    mcc_diff = mcc_attribute - mcc_mp,
    log_loss_diff = log_loss_attribute - log_loss_mp) %>%
  ungroup()
```

```{r, fig.width=10, fig.height=6, dpi=300}
ggplot(attr_mp, aes(x = replicate, y = accuracy_diff)) +
  geom_point(shape = 21, size = 2, aes(fill = data_source)) +
  geom_hline(yintercept = 0,
             color = "black", linetype = "dashed", linewidth = 0.8) +
  scale_fill_manual(
    name = "Accuracy Difference",
    values = c("CDM" = "darkolivegreen1", "IRT" = "deeppink2")
  ) +
  scale_x_continuous(name = "Replication", breaks = 1:10) +
  ylab("Accuracy Difference") +
  ggtitle("Accuracy Difference by Attribute - MP") +
  theme_bw() +
  facet_grid(structure + dim ~ correlation)

ggplot(attr_mp, aes(x = replicate, y = log_loss_diff)) +
  geom_point(shape = 21, size = 2, aes(fill = data_source)) +
  geom_hline(yintercept = 0,
             color = "black", linetype = "dashed", linewidth = 0.8) +
  scale_fill_manual(
    name = "Log Loss Difference",
    values = c("CDM" = "darkolivegreen1", "IRT" = "deeppink2")
  ) +
  scale_x_continuous(name = "Replication", breaks = 1:10) +
  ylab("Log Loss Difference") +
  ggtitle("Log Loss Difference by Attribute - MP") +
  theme_bw() +
  facet_grid(structure + dim ~ correlation) 
```























